A weakly structured stem for human origins in Africa

### 2022-09-01 paper.tex
more minor edits
added Georgette in acknowledgements

By contrast, our inferred models paint a more consistent picture of the late
Middle Stone Age as a critical period of change, assuming that estimates from
the recombination clock accurately relate to geological chronologies
(SI Section~\ref{sec:supp-mutation}).
During the Middle Stone Age, the multiple merger model indicates three
major stem lineages in Africa, tentatively assigned to southern (Stem 1S),
eastern (Stem 1E) and western/central Africa (Stem 2). While the length of
isolation among the stems is variable across model fits, models with a period of 
divergence, isolation and then a merger event (i.e. a ``reticulation''
out-performed models with bifurcating divergence and continuous gene flow.

@@

We are grateful for the DNA contribution from each participant which enabled
this study; in particular we wish to highlight the generous participation of
the Richtersveld Nama community in South Africa and help from local research
assistants Willem DeKlerk and Hendrik Kaimann. Additional assistance and
community engagement was conducted by Justin Myrick, Chris Gignoux, Caitlen
Uren and Cedric Werely. We thank the African Genome Diversity Project for data
generation, including Tommy Carensten, Deepti Gurdasani, and Manj Sandhu. We
thank Luke Anderson-Trocm\'e for assistance in creating the map in
Figure~\ref{fig:diversity}. This research was supported by an NIH grant R35GM133531 (to
BMH). The content is solely the responsibility of the authors and does
not necessarily represent the official views of the National Institutes of
Health.

>>>>

By contrast, our inferred models paint a more consistent picture of the late
Middle Stone Age as a critical period of change, assuming that estimates from
the recombination clock accurately relate to geological chronologies
(SI Section~\ref{sec:supp-mutation}).
During the Middle Stone Age, the multiple merger model indicates three
major stem lineages in Africa, tentatively assigned to southern (Stem 1S),
eastern (Stem 1E) and western/central Africa (Stem 2). While the length of
isolation among the stems is variable across model fits, models with a period of 
divergence, isolation and then a merger event (i.e. a ``reticulation'')
out-performed models with bifurcating divergence and continuous gene flow.

@@

We are grateful for the DNA contribution from each participant which enabled
this study; in particular we wish to highlight the generous participation of
the Richtersveld Nama community in South Africa and help from local research
assistants Willem DeKlerk and Hendrik Kaimann. Additional assistance and
community engagement was conducted by Justin Myrick, Chris Gignoux, Caitlen
Uren and Cedric Werely. We thank the African Genome Diversity Project for data
generation, including Tommy Carensten, Deepti Gurdasani, and Manj Sandhu. We
thank Luke Anderson-Trocm\'e and Georgette Femerling for assistance in creating the map in
Figure~\ref{fig:diversity} and Figure \ref{fig:toymodels}, respectively. This research was supported by an NIH grant R35GM133531 (to
BMH). The content is solely the responsibility of the authors and does
not necessarily represent the official views of the National Institutes of
Health.
